Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational approach

### 2016-09-14 ms_pre_post.Rmd
Clean comments, add subtitle

---
title: "Evolution of precopulatory and postcopulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

```{r, echo=FALSE}
figscale <- 1.0; # Change the scale of a figure
```

@@

<!--- TODO: GENERAL PARAGRAPH INTRODUCING THE SCOPE AND KEY QUESTION --->
Reproductive decisions expressed among interacting individuals directly affect individual fitnesses and the genotypes and viabilities of resulting offspring, and thereby profoundly affect evolutionary dynamics through the differential inheritence of alleles over generations [@Pizzari2012; @Wright2013; @Wild2014; @Liao2015; @Reid2016]. For example, how frequently a focal female mates is a reproductive decision that can directly and indirectly affect her fitness, and potentially alter the distribution of genotypes and patterns of inheritence within a population [e.g., @Cornell2007; @Diaz2010; @Okada2015; @Roberts2015a; @Worthington2016]. Her mating frequency and expression of mate choice might be affected by the genetic quality or complementarity of available potential mates, and her choosiness might additionally be expressed after mating has taken place through post-copulatory mechanisms [@Parker1983; @Parker2006; @Kuijper2012]. Female choice might therefore be realised by evolution of multiple, potentially interdependent, behavioural and physiological phenotypes operating in tandem. However, no general modelling framework predicts the conditions under which female mating frequency, pre-copulatory choice, and post-copulatory choice evolve together, meaning that there are no formal hypotheses for predicting the relevance of these reproductive decisions across different biological systems.

<!--- TODO: PARAGRAGH ON INBREEDING, PRE-COP, AND POST-COP AVOIDANCE --->
Inbreeding, defined as breeding between relatives, directly affects inheritence of identical-by-descent alleles and the viability of inbred offspring, meaning that the decision to prefer, avoid, or tolerate (i.e., mate randomly with respect to relatedness) inbreeding is predicted to have profound effects on individual fitness [@Parker2006; @Kokko2006; @Szulkin2012; @Duthie2015a]. Viability of inbred offspring typically decreases with increasing relatedness between parents [termed 'inbreeding depression'; @Keller2002; @Charlesworth2009]. If inbreeding depression is sufficiently weak, selection for female inbreeding preference is predicted because inbreeding increases the reproductive success of a focal female's male relative, thereby increasing the number of identical-by-descent allele copies present in her offspring [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, if inbreeding depression is moderate to strong, selection for female inbreeding avoidance is predicted [Duthie and Reid *in press*; @Duthie]. Female inbreeding strategy can be enacted before or after copulation occurs [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, models of inbreeding do not consider pre-copulatory inbreeding strategy (hereafter 'inbreeding mating strategy') and post-copulatory inbreeding strategy (hereafter 'inbreeding fertilisation strategy') jointly, meaning that it is unclear when selection for either or both strategies is predicted in empirical systems.

<!--- such as when focal females mate multiply (i.e., polyandry) to avoid inbreeding  [e.g., @Mays2008; @Brouwer2011; @Kingma2013; @Duthie] --->

<!--- TODO: PARAGRAGH ON EVOLUTION OF POLYANDRY (H1) --->
Critically, evolution of inbreeding fertilisation strategy is precluded if females only mate with one male, meaning that evolution of polyandry (defined as female multiple mating within a single reproductive bout) is intrinsically tied to female inbreeding fertilisation strategy. Although polyandry is commonplace in wild populations, the regularity with which it is observed in the wild remains theoretically puzzling because engaging in polyandry does not necessarily increase the number of offspring that a female produces, meaning that the fitness benefits associated with polyandry are not always clear [@Parker2013; @Taylor2014]. It has been hypothesised that females can acheive a net fitness benefit if polyandry provides them with additional opportunities to adjust their inbreeding through copulations with additional males [e.g., @Tregenza2002; @Akcay2007; @Lehtonen2015]. @Duthie showed that selection for polyandry to facilitate inbreeding avoidance is sometimes predicted, but only under highly constrained conditions in which very few males are available during a female's initial mate choice but many males are available during a female's additional mate choice(s), or when polyandry is conditionally expressed based on the degree to which a focal female engaged in inbreeding given her initially chosen mate. However, @Duthie assumed that polyandrous females had no control over the assignment of paternity to their offspring. If after engaging in polyandry, females could bias paternity toward non-relatives, then selection for polyandry to avoid inbreeding might be more widely predicted [@Zeh1997]. However, no models quantify the degree to which expression of biased paternity resulting from evolving inbreeding fertilisation strategy can feedback to drive evolution of polyandry in populations of potentially inbreeding females and males with biologically realistic distributions of relatedness.

<!--- TODO: PARAGRAGH ON IDEA OF PRE VS POST COPULATORY INBREEDING STRATEGY --->
Evolution of inbreeding fertilisation strategy might increase selection for polyandry, but evolutionary feedbacks between inbreeding fertilisation strategy and polyandry might also be contingent upon evolution of inbreeding mating strategy, thereby leading to complex evolutionary dynamics between inbreeding strategies and polyandry. For example, if females can acheive their desired degree of inbreeding through mate choice alone, then subsequent polyandry and inbreeding fertilisation strategy might be superfluous for increasing female fitness. Alternatively, if females are often unsuccessful at realising their inbreeding mating strategy due to factors such as sexual conflict over mating and resulting male coercion [e.g., @Parker2006], limited kin recognition amongst potential mates [e.g., @Penn2010], or simply the inherent stochasticity associated with mate encounters or expression of mate choice based on inbreeding [e.g., @Kokko2006; @Duthie], then selection on inbreeding fertilisation strategy might occur to reinforce inbreeding mating strategy. In empirical systems where inbreeding depression is severe and inbreeding avoidance is therefore beneficial, expression of inbreeding mating strategy and inbreeding fertilisation strategy have been observed across diverse taxa [e.g., @Gasparini2011; @Tuni2013; @Firman2015; @Fischer2015; @Daniel2015]. However, no models predict *a priori* which of these inbreeding strategies should be observed when inbreeding depression is sufficiently severe to make inbreeding avoidance adaptive.

<!--- TODO: PARAGRAGH ON COSTS OF POLYANDRY, MATING, AND FERTILISATION STRATEGIES --->
Evolution of polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy will ultimately be determined by the, potentially interacting, fitness benefits and costs of each phenotype. Given severe inbreeding depression, phenotypes will increase a focal female's fitness if they cause her to avoid breeding with close relatives [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, evolution of phenotypes will also be directly affected by negative selection (i.e., costs) associated with the tendency to engage in polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy. For example, females with a tendency to engage in polyandry might pay a cost associated with searching or making themselves available to mates, thereby increasing their own risk of predation [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that are particularly choosy in mating interactions (e.g., especially reluctant to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that are choosy after mating has occurred might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. Overall, to predict evolution of polyandry and inbreeding strategies, it is therefore necessary to model costs associated with polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy.

<!--- Mate searching: Lehmann2013, Kokko2006, Jamieson2009. Kin discrimination: Waser1986. Increased predation risk associated with mate search: Rowe1988, Ronkainen1994, Koga1998. Increased risk associated with expression of mate choice (e.g., avoiding male coercion): Rowe1994. Risk of not mating due to choosiness: Kokko2013

<!--- TODO: PARAGRAGH ON ORDER OF INTERACTIONS --->
Selection for inbreeding mating strategy versus inbreeding fertilisation strategy and associated polyandry might also depend evolutionary constraints, or on existing phenotypes in the population. For example, if phenotypes biasing fertilisation to avoid inbreeding are precluded by evolutionary constraints [@Bradshaw1991; @Futuyma2010], then inbreeding avoidance might only be possible before copulation occurs, perhaps strengthening selection for inbreeding mating strategy and weakening selection for polyandry. Alternatively, evolution of female inbreeding mating strategy might be precluded due to severely limited kin recognition or the inability to resist male coercion given sexual conflict [@Arnqvist2005a; @Parker2006; @Penn2010]. In such conditions, females might only be able to express choice through inbreeding fertilisation strategy, perhaps strengthening selection for inbreeding avoidance through polyandry and post-copulatory mechanisms. Selection for inbreeding mating or fertilisation strategy might also be weak or negligible if the strategy is redundant due to the presence of an already established adaptive inbreeding strategy (e.g., if females already successfully avoid inbreeding during mate encounters, then they might be unable to increase their fitness further by biasing fertilisation after mating has occurred). It is therefore necessary to consider evolution of inbreeding mating strategy and inbreeding fertilisation strategy when each evolves in a population where its complementary inbreeding strategy is evolutionarily constrained or already fixed as an adaptation.

<!--- TODO: OVERVIEW OF THE POINT OF THE MODELLING --->
Here we apply a model of female and male reproductive interactions and associated inbreeding depression in offspring viability to quantify evolution of three separate phenotypes including female tendency for polyandry, female pre-copulatory inbreeding strategy (mating strategy), and female post-copulatory inbreeding strategy (fertilisation strategy). We use individual-based modelling and *in silico* experiments to test the hypothesis that, given severe inbreeding depression in offspring viability, evolution of inbreeding avoidance through fertilisation bias causes selection for evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution of each such that that inbreeding avoidance through mate choice but not fertilisation bias will evolve when inbreeding mating strategy is less costly than the combined cost of inbreeding fertilisation strategy and tendency for polyandry, while inbreeding avoidance through fertilisation bias but not mate choice will evolve when inbreeding mating strategy is more costly than the combined costs of inbreeding fertilisation strategy and tendency for polyandry. Finally, we test the hypothesis that constraints on evolution of one inbreeding strategy (e.g., inbreeding mating strategy) will increase the strength of selection on evolution of a complementary inbreeding strategy (e.g., inbreeding fertilisation strategy), and that evolution of one inbreeding strategy will be precluded if it is introduced into a population that is already fixed for a complementary inbreeding strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and strong inbreeding depression in offspring viability.

@@

<!--- Change mating cost to an increased risk of mating failure ---> 
Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilistation strategy ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilisation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of inbreeding mating strategy and inbreeding fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

<!--- Figure 2 shows an equivalent figure with changes in allele values of tendency for polyandry (red lines and shading) and inbreeding mating strategy (blue lines and shading) only. However, in Figure 2, a horizontal black line where mean allele value of inbreeding fertilisation strategy equals zero shows that these allele values have been fixed. Figure 2 has only four panels because costs of inbreeding fertilisation strategy is irrelevant when inbreeding fertilisation strategy alleles have no effect and mating is random (i.e., $F_{p}=0$). --->

@@

<!--- Figure 1 will show a spread of all the results (evo combs02 B3.eps), and be contrasted with the four panel Figure 2 showing results when $F_{p}$ is knocked out (evo combs02 B3 ko F.eps). --->

@@

<!--- Interestingly, a high cost of only tendency for polyandry (Figure 1C) or only inbreeding fertilisation strategy (Figure 1D) shows how these phenotypes are evolutionarily tied such that a cost of one phenotype can feed back to affect evolution of the other phenotype. For example, Figure 1C shows simulations in which inbreeding fertilisation strategy is not costly, causing $F_{a}$ values to become more negative (stronger inbreeding avoidance) than when both inbreeding fertilisation strategy and tendency for polyandry are costly (Figure 1G). However, $F_{a}$ values are not as negative as they would be if neither tendency for polyandry or inbreeding fertilisations trategy were costly (Figure 1A) because expression of fertilisation bias was hindered by costs of polyandry and an associated reduction in evolved mating frequency. Similarly, Figure 1D shows simulations in which only inbreeding fertilisation strategy was costly, causing less strong inbreeding avoidance through fertilisation bias. Consequently, tendency for polyandry values were lower than when inbreeding fertilisation strategy was not costly (Figure 1A) because polyandry was less beneficial when it did not result in strong inbreeding avoidance through fertilisation bias. --->

@@

<!--- Figure 3 will show constraints on $M_{p}$ affecting $F_{p}$, while Figure 2 will be used again to show constraints on $F_{p}$ affecting $M_{p}$. --->

@@

<!--- Figure 4 will show evolution of $M_{p}$ given pre-existing post-copulatory inbreeding avoidance, and Figure 5 will show evolution of $P_{p}$ and $F_{p}$ given pre-existing pre-copulatory inbreeding avoidance. --->

<!--- Get in there that the replicate simulations don't show evidence of alternative states -- one thing happens or does not, though there is some correlation among traits in replicates --->

@@

<!--- (1) selection for polyandry is strengthened given adaptive inbreeding avoidance through fertilisation bias, (2) selection for pre- versus post-copulatory inbreeding avoidance will be strong for the least costly, and negligible for the most costly, mechanism of inbreeding avoidance, (3) selection for pre- or post-copulatory inbreeding avoidance will be strengthened if evolutionary constraints preclude inbreeding avoidance in the complementary mechanism, and (4) selection for pre- versus post-copulatory inbreeding avoidance will be weakened if a complementary adaptive mechanism of inbreeding avoidance is already fixed in a population.--->

@@

<!--- Costs of polyandry including threshold trait -> polyandry affected by cost of inbreeding fertilisation strategy -> Polyandry indirectly affected by cost of inbreeding mating strategy through selection on post-copulatory inbreeding avoidance. Cost of inbreeding strategies mostly independent of one another and cost of polyandry.  --->

<!--- For example, females with a tendency to engage in polyandry might pay a cost associated with searching or making themselves available to mates, thereby increasing their own risk of predation [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that are particularly choosy in mating interactions (e.g., especially reluctant to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that are choosy after mating has occurred might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b] --->

>>>>

---
title: "Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational approach"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

@@

Reproductive decisions expressed among interacting individuals directly affect individual fitnesses and the genotypes and viabilities of resulting offspring, and thereby profoundly affect evolutionary dynamics through the differential inheritence of alleles over generations [@Pizzari2012; @Wright2013; @Wild2014; @Liao2015; @Reid2016]. For example, how frequently a focal female mates is a reproductive decision that can directly and indirectly affect her fitness, and potentially alter the distribution of genotypes and patterns of inheritence within a population [e.g., @Cornell2007; @Diaz2010; @Okada2015; @Roberts2015a; @Worthington2016]. Her mating frequency and expression of mate choice might be affected by the genetic quality or complementarity of available potential mates, and her choosiness might additionally be expressed after mating has taken place through post-copulatory mechanisms [@Parker1983; @Parker2006; @Kuijper2012]. Female choice might therefore be realised by evolution of multiple, potentially interdependent, behavioural and physiological phenotypes operating in tandem. However, no general modelling framework predicts the conditions under which female mating frequency, pre-copulatory choice, and post-copulatory choice evolve together, meaning that there are no formal hypotheses for predicting the relevance of these reproductive decisions across different biological systems.

Inbreeding, defined as breeding between relatives, directly affects inheritence of identical-by-descent alleles and the viability of inbred offspring, meaning that the decision to prefer, avoid, or tolerate (i.e., mate randomly with respect to relatedness) inbreeding is predicted to have profound effects on individual fitness [@Parker2006; @Kokko2006; @Szulkin2012; @Duthie2015a]. Viability of inbred offspring typically decreases with increasing relatedness between parents [termed 'inbreeding depression'; @Keller2002; @Charlesworth2009]. If inbreeding depression is sufficiently weak, selection for female inbreeding preference is predicted because inbreeding increases the reproductive success of a focal female's male relative, thereby increasing the number of identical-by-descent allele copies present in her offspring [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, if inbreeding depression is moderate to strong, selection for female inbreeding avoidance is predicted [Duthie and Reid *in press*; @Duthie]. Female inbreeding strategy can be enacted before or after copulation occurs [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, models of inbreeding do not consider pre-copulatory inbreeding strategy (hereafter 'inbreeding mating strategy') and post-copulatory inbreeding strategy (hereafter 'inbreeding fertilisation strategy') jointly, meaning that it is unclear when selection for either or both strategies is predicted in empirical systems.

Critically, evolution of inbreeding fertilisation strategy is precluded if females only mate with one male, meaning that evolution of polyandry (defined as female multiple mating within a single reproductive bout) is intrinsically tied to female inbreeding fertilisation strategy. Although polyandry is commonplace in wild populations, the regularity with which it is observed in the wild remains theoretically puzzling because engaging in polyandry does not necessarily increase the number of offspring that a female produces, meaning that the fitness benefits associated with polyandry are not always clear [@Parker2013; @Taylor2014]. It has been hypothesised that females can acheive a net fitness benefit if polyandry provides them with additional opportunities to adjust their inbreeding through copulations with additional males [e.g., @Tregenza2002; @Akcay2007; @Lehtonen2015]. @Duthie showed that selection for polyandry to facilitate inbreeding avoidance is sometimes predicted, but only under highly constrained conditions in which very few males are available during a female's initial mate choice but many males are available during a female's additional mate choice(s), or when polyandry is conditionally expressed based on the degree to which a focal female engaged in inbreeding given her initially chosen mate. However, @Duthie assumed that polyandrous females had no control over the assignment of paternity to their offspring. If after engaging in polyandry, females could bias paternity toward non-relatives, then selection for polyandry to avoid inbreeding might be more widely predicted [@Zeh1997]. However, no models quantify the degree to which expression of biased paternity resulting from evolving inbreeding fertilisation strategy can feedback to drive evolution of polyandry in populations of potentially inbreeding females and males with biologically realistic distributions of relatedness.

Evolution of inbreeding fertilisation strategy might increase selection for polyandry, but evolutionary feedbacks between inbreeding fertilisation strategy and polyandry might also be contingent upon evolution of inbreeding mating strategy, thereby leading to complex evolutionary dynamics between inbreeding strategies and polyandry. For example, if females can acheive their desired degree of inbreeding through mate choice alone, then subsequent polyandry and inbreeding fertilisation strategy might be superfluous for increasing female fitness. Alternatively, if females are often unsuccessful at realising their inbreeding mating strategy due to factors such as sexual conflict over mating and resulting male coercion [e.g., @Parker2006], limited kin recognition amongst potential mates [e.g., @Penn2010], or simply the inherent stochasticity associated with mate encounters or expression of mate choice based on inbreeding [e.g., @Kokko2006; @Duthie], then selection on inbreeding fertilisation strategy might occur to reinforce inbreeding mating strategy. In empirical systems where inbreeding depression is severe and inbreeding avoidance is therefore beneficial, expression of inbreeding mating strategy and inbreeding fertilisation strategy have been observed across diverse taxa [e.g., @Gasparini2011; @Tuni2013; @Firman2015; @Fischer2015; @Daniel2015]. However, no models predict *a priori* which of these inbreeding strategies should be observed when inbreeding depression is sufficiently severe to make inbreeding avoidance adaptive.

Evolution of polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy will ultimately be determined by the, potentially interacting, fitness benefits and costs of each phenotype. Given severe inbreeding depression, phenotypes will increase a focal female's fitness if they cause her to avoid breeding with close relatives [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, evolution of phenotypes will also be directly affected by negative selection (i.e., costs) associated with the tendency to engage in polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy. For example, females with a tendency to engage in polyandry might pay a cost associated with searching or making themselves available to mates, thereby increasing their own risk of predation [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that are particularly choosy in mating interactions (e.g., especially reluctant to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that are choosy after mating has occurred might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. Overall, to predict evolution of polyandry and inbreeding strategies, it is therefore necessary to model costs associated with polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy.

Selection for inbreeding mating strategy versus inbreeding fertilisation strategy and associated polyandry might also depend evolutionary constraints, or on existing phenotypes in the population. For example, if phenotypes biasing fertilisation to avoid inbreeding are precluded by evolutionary constraints [@Bradshaw1991; @Futuyma2010], then inbreeding avoidance might only be possible before copulation occurs, perhaps strengthening selection for inbreeding mating strategy and weakening selection for polyandry. Alternatively, evolution of female inbreeding mating strategy might be precluded due to severely limited kin recognition or the inability to resist male coercion given sexual conflict [@Arnqvist2005a; @Parker2006; @Penn2010]. In such conditions, females might only be able to express choice through inbreeding fertilisation strategy, perhaps strengthening selection for inbreeding avoidance through polyandry and post-copulatory mechanisms. Selection for inbreeding mating or fertilisation strategy might also be weak or negligible if the strategy is redundant due to the presence of an already established adaptive inbreeding strategy (e.g., if females already successfully avoid inbreeding during mate encounters, then they might be unable to increase their fitness further by biasing fertilisation after mating has occurred). It is therefore necessary to consider evolution of inbreeding mating strategy and inbreeding fertilisation strategy when each evolves in a population where its complementary inbreeding strategy is evolutionarily constrained or already fixed as an adaptation.
Here we apply a model of female and male reproductive interactions and associated inbreeding depression in offspring viability to quantify evolution of three separate phenotypes including female tendency for polyandry, female pre-copulatory inbreeding strategy (mating strategy), and female post-copulatory inbreeding strategy (fertilisation strategy). We use individual-based modelling and *in silico* experiments to test the hypothesis that, given severe inbreeding depression in offspring viability, evolution of inbreeding avoidance through fertilisation bias causes selection for evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution of each such that that inbreeding avoidance through mate choice but not fertilisation bias will evolve when inbreeding mating strategy is less costly than the combined cost of inbreeding fertilisation strategy and tendency for polyandry, while inbreeding avoidance through fertilisation bias but not mate choice will evolve when inbreeding mating strategy is more costly than the combined costs of inbreeding fertilisation strategy and tendency for polyandry. Finally, we test the hypothesis that constraints on evolution of one inbreeding strategy (e.g., inbreeding mating strategy) will increase the strength of selection on evolution of a complementary inbreeding strategy (e.g., inbreeding fertilisation strategy), and that evolution of one inbreeding strategy will be precluded if it is introduced into a population that is already fixed for a complementary inbreeding strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and strong inbreeding depression in offspring viability.

@@

Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilistation strategy ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilisation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of inbreeding mating strategy and inbreeding fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

@@

@@

@@

@@

@@

@@
